---
title: "9762 analysis"
author: "Andrew Portuguese"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  distill::distill_article:
    code_folding: true
    toc: true
    toc_float: true
    self_contained: true
---

# Setup and data wrangling

## Load necessary packages
```{r, warning=FALSE}
library(gtsummary)
library(survival)
library(readxl)
library(dplyr)
library(lubridate)
library(readr)
library(survminer)
library(ggsurvfit)
library(rms)
library(gt)
library(bstfun)
```

## Read in the data
```{r}
db1 = read_excel(path = "9762 data for analysis 1.23.24.xlsx", sheet = "Survival")

dataset <- db1 %>%
  mutate(
    dx_dt = ymd(dx_dt),
    infuse_dt = ymd(infuse_dt),
    birthdate = ymd(birthdate),
    dt_relapse = ymd(dt_relapse),
    Death.or.DLC = ymd(Death.or.DLC),
    Death.or.relapse.or.DLC = ymd(Death.or.relapse.or.DLC),
    Days.to.death.or.DLC = Death.or.DLC - infuse_dt,
    Days.to.death.or.relapse.or.DLC = Death.or.relapse.or.DLC - infuse_dt,
    Days.from.dx.to.infusion = infuse_dt - dx_dt,
    Age = (infuse_dt-birthdate)/365
  )
```


# Data tables

## Table 1 reproduction
```{r}
theme_gtsummary_compact()


dataset %>%
  transmute(
    "Age (years)" = Age,
    "Female sex" = ifelse(sex == "f", 1, 0),
    "Race" = recode(race, "asian" = "Asian", "white" = "White"),
    "ECOG performance status" = ecog,
    "Median time from diagnosis to CAR T cell therapy" = Days.from.dx.to.infusion/365,
    "Disease subtype" = factor(subtype_category, levels = c("IgA", "IgG", "IgM", "Light chain", "Oligosecretory", "Primary plasma cell leukemia")),
    "ISS disease stage at diagnosis" = iss,
    "Revised ISS disease stage at diagnosis" = riss,
    "ISS disease stage at enrollment" = iss_enrollment,
    "Revised ISS disease stage at enrollment" = riss_enrollment,
    #"High-risk cytogenetic abnormality at any time" = HRCA,
    "t(4;14) any" = t414_any,
    "t(14;16) any" = t1416_any,
    "del(17p) any" = del17p_any,
    "amp(1q) any" = amp1q_any,
    "High bone marrow tumor burden" = ifelse(High.tumor.burden == 1, ">30%", "10-30%"),
    "Extramedullary disease" = extramedullary,
    "Circulating plasma cells at screening" = cpcs,
    "Tumor BCMA expression (ABC)" = bcma_scrn,
    "Number of previous antimyeoma regimens" = factor(case_when(
      no_priortx <=6 ~ "4-6",
      no_priortx >6 & no_priortx <=9 ~ "7-9",
      no_priortx >9 ~ "10 or more"), levels = c("4-6", "7-9", "10 or more") ),
    "Previous autologous HCT" = factor(case_when(
      no_auto == 0 ~ "0",
      no_auto == 1 ~ "1",
      no_auto >1 ~ ">1"), levels = c("0","1",">1") ),
    "Previous allogeneic HCT" = allo
  ) %>%
  
  tbl_summary(
    missing = "ifany",
    statistic = list(
      all_continuous() ~ "{median} ({min} to {max})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  bold_labels() %>%
  add_variable_grouping(
    "High-risk cytogenetic abnormality at any time" = c("t(4;14) any", "t(14;16) any", "del(17p) any", "amp(1q) any")
  )

```



# Survival analysis
## Length of follow-up
```{r}
data1 = dataset %>%
  mutate(
    Reverse_death = ifelse(death == 1, 0,1)
  )

#quantile(prodlim(Hist(time = Real.days.DLC/30, Real.death )~1, data = data1, reverse = TRUE ) )


reverse_km_OS <- survfit(Surv(Days.to.death.or.DLC/30, Reverse_death) ~ 1, data1)
reverse_km_OS
```
## Create KM for OS and PFS
```{r, warning=FALSE, message=FALSE}

km_OS <- survfit(Surv(Days.to.death.or.DLC/30, death) ~ 1, data = dataset)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
) 

OS$plot <- OS$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_OS$median + 0.5, y = 0.54, label = paste(signif(med_OS$median, 2), "(95% CI, ", signif(med_OS$lower, 2), "to", signif(med_OS$upper, 2),")"), hjust = 0,  size = 5)
OS
km_OS
summary(km_OS,times=c(12))

km_PFS <- survfit(Surv(Days.to.death.or.relapse.or.DLC/30, relapse) ~ 1, data = dataset)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS$plot <- PFS$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_PFS$median + 0.5, y = 0.54, label = paste(signif(med_PFS$median, 2), "(95% CI, ", signif(med_PFS$lower, 2), "to", signif(med_PFS$upper, 2),")"), hjust = 0,  size = 5)
PFS
km_PFS
summary(km_PFS,times=c(12))
```

## PFS and OS on the same plot
```{r}
OS_EFS <- ggsurvplot_combine(list(km_PFS, km_OS),
           data=db,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.3,
           tables.theme = theme_cleantable(),
           surv.median.line = "hv", # Specify median survival
           ggtheme = theme_classic()+
            theme(
              axis.text.x = element_text(size = 12),  # Increase x-axis label font size
              axis.text.y = element_text(size = 12),  # Increase y-axis label font size
              axis.title.x = element_text(size = 14),  # Increase x-axis title font size
              axis.title.y = element_text(size = 14),  # Increase y-axis title font size
              plot.title = element_text(size = 12),    # Increase plot title font size
              legend.text = element_text(size = 12),    # Increase legend text font size
              legend.title = element_text(size = 12)    # Increase legend title font size
            ),
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "PFS & OS (%)",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           #fun=formula,
           legend.labs = c("PFS","OS"),
           surv.scale = "percent",
           censor=FALSE
           )

OS_EFS$plot <- OS_EFS$plot + 
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") + 
  geom_vline(xintercept = median(km_OS)[1], linetype = "dashed", color = "grey") + 
  geom_vline(xintercept = median(km_PFS)[1], linetype = "dashed", color = "grey") + 
  annotate("text", x = med_OS$median + 0.5, y = 0.85, label = paste(signif(med_OS$median, 2), "(", signif(med_OS$lower, 2), "-", signif(med_OS$upper, 2),")"), hjust = 0,  size = 5) +
  annotate("text", x = med_PFS$median + 0.5, y = 0.95, label = paste(signif(med_PFS$median, 2), "(", signif(med_PFS$lower, 2), "-", signif(med_PFS$upper, 2),")"), hjust = 0,  size = 5)

OS_EFS
```





## Create KM for DOR
```{r, warning=FALSE, message=FALSE}


km_DOR <- survfit(Surv((Days.to.death.or.relapse.or.DLC-28)/30, relapse) ~ 1, data = dataset)
med_DOR <- surv_median(km_DOR)

DOR <- ggsurvplot(km_DOR,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Duration of response",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

DOR$plot <- DOR$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_DOR$median + 0.5, y = 0.54, label = paste(signif(med_DOR$median, 2), "(95% CI, ", signif(med_DOR$lower, 2), "to", signif(med_DOR$upper, 2),")"), hjust = 0,  size = 5)
DOR
km_DOR
summary(km_DOR,times=c(12))
```

# Stratified survival analysis
## Create KM for OS and PFS, stratify by HRCA
```{r, warning=FALSE, message=FALSE}

km_OS <- survfit(Surv(Days.to.death.or.DLC/30, death) ~ HRCA, data = dataset)
km_OS

med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black"),
           legend.labs = c("Standard","High risk")
) 

#OS$plot <- OS$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_OS$median + 0.5, y = 0.54, label = paste("Median =", signif(med_OS$median, 2), "(95% CI, ", signif(med_OS$lower, 2), "to", signif(med_OS$upper, 2),")"), hjust = 0,  size = 5)

OS

km_PFS <- survfit(Surv(Days.to.death.or.relapse.or.DLC/30, relapse) ~ HRCA, data = dataset)
km_PFS

med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black"),
           legend.labs = c("Standard","High risk")
)

#PFS$plot <- PFS$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_PFS$median + 0.5, y = 0.54, label = paste("Median =", signif(med_PFS$median, 2), "(95% CI, ", signif(med_PFS$lower, 2), "to", signif(med_PFS$upper, 2),")"), hjust = 0,  size = 5)

PFS
```

## KM of OS & PFS, stratified by cell dose
```{r}
km_PFS <- survfit(Surv(Days.to.death.or.relapse.or.DLC/30, relapse) ~ intended_cell_dose, data = dataset)
km_PFS

med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.4,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black"),
           #legend.labs = c("Standard","High risk")
)

PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ intended_cell_dose, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.4,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS

```

## KM of PFS, dose level 1 vs 2-4
```{r}

data1 <- dataset %>%
  mutate(
    dose_level = ifelse(intended_cell_dose == "5e+07","1","2-4")
  )

km_PFS <- survfit(Surv(Days.to.death.or.relapse.or.DLC/30, relapse) ~ dose_level, data = data1)
km_PFS

med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("1","2-4"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black"),
           #legend.labs = c("Standard","High risk")
)

#PFS$plot <- PFS$plot + geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey") +  annotate("text", x = med_PFS$median + 0.5, y = 0.54, label = paste("Median =", signif(med_PFS$median, 2), "(95% CI, ", signif(med_PFS$lower, 2), "to", signif(med_PFS$upper, 2),")"), hjust = 0,  size = 5)

PFS

```

## Create KM for OS, stratified by sCR/CR vs VGPR/PR
```{r, warning=FALSE, message=FALSE}

data1 <- dataset %>%
  mutate(
    response.category = recode(best_response, "scr" = "sCR/CR", "cr" = "sCR/CR", "vgpr" = "VGPR/PR", "pr"="VGPR/PR")
  )

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ response.category, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("sCR/CR","VGPR/PR"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS


km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC-28)/30, relapse) ~ response.category, data = data1)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("sCR/CR","VGPR/PR"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS
```

## Create KM for OS & PFS, stratified by treatment lines <8 vs >=8
```{r, warning=FALSE, message=FALSE}

data1 <- dataset %>%
  mutate(
    prior.lines.8 = ifelse(no_priortx >=8, 1,0)
  )

km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ prior.lines.8, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("<8","\u22658"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS


km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ prior.lines.8, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("<8","\u22658"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by hx of prior allo
```{r, warning=FALSE, message=FALSE}

data1 <- dataset %>%
  mutate(
    prior.lines.8 = ifelse(no_priortx >=8, 1,0)
  )

km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ allo, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS
summary(km_PFS,times=c(12))


km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ allo, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
summary(km_OS,times=c(12))
```

## Create KM for OS & PFS, stratified by receipt of prior BCMA-directed therapy
```{r, warning=FALSE, message=FALSE}

data1 <- dataset %>%
  mutate(
    prior.lines.8 = ifelse(no_priortx >=8, 1,0)
  )

km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ any_bcma_tx, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS
summary(km_PFS,times=c(12))


km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ any_bcma_tx, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
summary(km_OS,times=c(12))
```

## Create KM for OS & PFS, stratified by triple refractory disease
```{r, warning=FALSE, message=FALSE}
km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ triple_ref, data = dataset)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ triple_ref, data = dataset)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by penta refractory disease
```{r, warning=FALSE, message=FALSE}
km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ pentaref, data = dataset)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ triple_ref, data = dataset)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by tumor burden (BMPCs <=30 vs >30)
```{r, warning=FALSE, message=FALSE}
km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ High.tumor.burden, data = dataset)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ High.tumor.burden, data = dataset)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by sBCMA at day 90 (< vs >= median)
```{r, warning=FALSE, message=FALSE}
data1 <- dataset %>%
  filter(!is.na(sBCMA.d90))%>%
  filter(Days.to.death.or.relapse.or.DLC>=90) %>%
  mutate(
    sBCMA.high = ifelse(sBCMA.d90 >= median(sBCMA.d90),1,0)
  )

median(data1$sBCMA.d90)


km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC-90)/30, relapse) ~ sBCMA.high, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC-90)/30, death) ~ sBCMA.high, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by sBCMA at day 60 (< vs >= median)
```{r, warning=FALSE, message=FALSE}
data1 <- dataset %>%
  filter(!is.na(sBCMA.d60))%>%
  filter(Days.to.death.or.relapse.or.DLC>=60) %>%
  mutate(
    sBCMA.high = ifelse(sBCMA.d60 >= median(sBCMA.d60),1,0)
  )

median(data1$sBCMA.d60)

km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC-60)/30, relapse) ~ sBCMA.high, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC-60)/30, death) ~ sBCMA.high, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

## Create KM for OS & PFS, stratified by BCMA ABC at screening
```{r, warning=FALSE, message=FALSE}
data1 <- dataset %>%
  filter(!is.na(bcma_scrn))%>%
  filter(Days.to.death.or.relapse.or.DLC>=60) %>%
  mutate(
    BCMA.high = ifelse(bcma_scrn >= median(bcma_scrn),1,0)
  )

median(data1$bcma_scrn)

km_PFS <- survfit(Surv((Days.to.death.or.relapse.or.DLC)/30, relapse) ~ BCMA.high, data = data1)
med_PFS <- surv_median(km_PFS)

PFS <- ggsurvplot(km_PFS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Progression-free survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

PFS
km_PFS

km_OS <- survfit(Surv((Days.to.death.or.DLC)/30, death) ~ BCMA.high, data = data1)
med_OS <- surv_median(km_OS)

OS <- ggsurvplot(km_OS,
           pval=TRUE,
           conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           fontsize = 6,
           risk.table.col = "strata", # Change risk table color by groups
           tables.height = 0.25,
           tables.theme = theme_cleantable(),
           #surv.median.line = "hv", # Specify median survival
           #ggtheme = theme_bw(), # Change ggplot2 theme
           #palette = c("#E7B800", "#2E9FDF"),
           xlab="Time (months)", ylab = "Overall survival",
           xlim = c(0,65), break.x.by = c(12),
           ylim = c(0,1), break.y.by = c(0.25),
           legend.labs = c("Low","High"),
           legend = "none",
           surv.scale="percent",
           font.main = c(16, "plain", "black"),
           font.x = c(16, "plain", "black"),
           font.y = c(16, "plain", "black"),
           font.caption = c(16, "plain", "black"),
           font.tickslab = c(16, "plain", "black")
)

OS
km_OS
```

# Regression analysis
## Cox regression
```{r}
theme_gtsummary_compact()
preds <- c("HRCA", "cnLOH", "response.category", "intended_cell_dose", "High.tumor.burden","any_bcma_tx","pentaref","triple_ref", "sBCMA.d90", "sBCMA.d60")

data1 <- dataset %>%
  mutate(
    response.category = recode(best_response, "scr" = "sCR/CR", "cr" = "sCR/CR", "vgpr" = "VGPR/PR", "pr"="VGPR/PR"),
    intended_cell_dose = log10(intended_cell_dose)
  )


uv_tab_OS <- tbl_uvregression(
  data1[c(preds)],
  method = coxph,
  y = Surv(data1$Days.to.death.or.DLC, data1$death),
  exponentiate = TRUE
) %>%
  sort_p()

uv_tab_PFS <- tbl_uvregression(
  data1[c(preds)],
  method = coxph,
  y = Surv(data1$Days.to.death.or.relapse.or.DLC, data1$relapse),
  exponentiate = TRUE
)


tbl_merge(list(uv_tab_OS, uv_tab_PFS), tab_spanner = c("**OS**", "**PFS**"))


```

# Scatterplot

## TTR by BCMA ABC at baseline in relapsed patients
```{r}
data1 <- dataset %>%
  filter(relapse == 1) %>%
  transmute(
    bcma_scrn = log(bcma_scrn),
    time.to.relapse = Days.to.death.or.relapse.or.DLC/30
  )

ggplot(data1, aes(x=time.to.relapse, y = bcma_scrn)) + 
  geom_point(shape = 16) + # Add points to the plot with shape 16 (i.e., basic circle)
  theme_classic() + # Use the minimal theme
  geom_smooth(aes(x = time.to.relapse, y=bcma_scrn), method = "lm", se = TRUE, color = "#104a8e", inherit.aes = FALSE, linetype = "solid", size = 0.5)+
  stat_cor(method = "pearson", label.x = 3, label.y = 9) +
  labs(x = "Time to relapse (months)", y = "log(BCMA ABC)")
  
```



# Probability plots
## Hazard of OS vs sBCMA d90
```{r}
data1 <- dataset %>%
  select(sBCMA.d90, Days.to.death.or.DLC, death) %>%
  filter(!is.na(sBCMA.d90))%>%
  mutate(
    sBCMA.high = ifelse(sBCMA.d90 >= median(sBCMA.d90),1,0)
  )



dd <- datadist(data1)
options(datadist='dd')

S <- Surv(data1$Days.to.death.or.DLC, data1$death)
f <- cph(S ~ sBCMA.d90, x=TRUE, y=TRUE,surv=TRUE,data=data1)

model <- Predict(f, sBCMA.d90, fun=function(x) exp(x) )

ggplot(as.data.frame(model),aes(x=sBCMA.d90, y=yhat)) + 
  geom_ribbon(data = model, aes(ymin=lower, ymax=upper), alpha=0.2, linetype=0) + 
  geom_line() + 
  theme_bw() + 
  labs(
    x = "sBCMA d90",
    y = "OS (hazard)") +
  theme(title = element_text(size=16,face="bold"),legend.title=element_blank())
```

## Hazard of OS vs screening BCMA ABC
```{r}
data1 <- dataset %>%
  select(bcma_scrn, Days.to.death.or.DLC, death)



dd <- datadist(data1)
options(datadist='dd')

S <- Surv(data1$Days.to.death.or.DLC, data1$death)
f <- cph(S ~ rcs(bcma_scrn,3), x=TRUE, y=TRUE,surv=TRUE,data=data1)

model <- Predict(f, bcma_scrn, fun=function(x) exp(x) )

ggplot(as.data.frame(model),aes(x=bcma_scrn, y=yhat)) + 
  geom_ribbon(data = model, aes(ymin=lower, ymax=upper), alpha=0.2, linetype=0) + 
  geom_line() + 
  theme_bw() + 
  labs(
    x = "sBCMA d90",
    y = "OS (hazard)") +
  theme(title = element_text(size=16,face="bold"),legend.title=element_blank())
```

